Towards a Wide-Coverage Grammar for Swedish\\ Using GF

### 2012-01-19 doc/report/report.tex
last report update befor submit

% check references! lncs is good maybe. lnai
\newpage
%\pagestyle{empty}
%\cleardoublepage
\tableofcontents
%
\addtocontents{toc}{\protect\thispagestyle{empty}}
%\pagestyle{empty}
%\cleardoublepage
\newpage
%\pagenumberingstyle{plain}
\setcounter{page}{1} %reset the page counter
\chapter{Introduction}
\thispagestyle{plain}
% more intro to gf. dependently typed...high-level formalism
Grammatical Framework \cite{gfbok} is a dependently typed grammar formalism.
%with facilities for multilinguality. 
It is based on Martin-Löf type theory which allows
reasoning within the programming language.
GF has strong support for multilinguality and has so far been used successfully for 
controlled languages \cite{cnl}. Recent experiments have  however showed
that it is also possible to use the framework for parsing free language \cite{patent}.
% >> although the success rate is going steeply down when using uncontrolled nl. Better with annotated.
% --> sentence about connection GF - grammar based parser.
%This project focuses on implementing a wide-coverage Swedish grammar, which
%describes a parser accepting all sentences that can be expressed by the grammar.
Parsing, that is the task of automatically identifying morphological and
syntactical structures, is receiving increasing interest, especially
considering the steadily growing amount of data available online. 
So far there is no freely available grammar-driven parser that gives a deep
analysis for Swedish. The fact that %the existing rule-based Swedish
a parser or grammar is not freely available does not only restrict its
usage but also its possibilities of being further developed. Our
goal is
to create an open-source Swedish grammar from which we derive a parser
accepting all sentences described by the given rules.
As a freely available resource, it can be continuously enhanced and
made use of in other projects.\\

@@

Apart from these restrictions, noun phrases containing reflexive pronouns
may be used as any other \verb-NP-. They may be conjoined 
(\ref{ex:ReflConj} a,b) and used with other determiners (\ref{ex:ReflPredet}).
In the standard GF analysis, which is preformed bottom-up starting from the
%TODO verify
POS-tags, information about semantic roles are given by the functions, not by the
categories. That is,
we know that the first argument of the function \verb-PredVP- acts as the subject.
The noun phrase itself does not carry information about its semantic role.
Until it is given as an argument to a clause level function, no difference
is made between subject and object noun phrases.
For this reason,
the formalization of reflexive pronouns required the use of a different analysis.
%and used as genitives (sentence \ref{ex:ReflGen}) as such.
%Could be combined with 'sig', a ReflNP without another noun.
In short, what is wanted can be summarized as follows:\\
%\begin{figure}[h]
\begin{enumerate}
\item{
Construct a noun phrase from a common noun phrase \\
katt $\rightarrow$ sin katt} % \\
\item{Modify it like a \verb|NP|\\
sina katter $\rightarrow$ alla sina katter}
\item{Use it only as object\\
sin katt $\rightarrow$ han såg sin katt }
\end{enumerate}
%\caption{Rules for noun phrases with reflexive possessive pronouns}
%\label{fig:reflrestrict}
%\end{figure}

>>>>

% check references! lncs is good maybe. lnai
\newpage
%\pagestyle{empty}
%\cleardoublepage
\tableofcontents
%
\addtocontents{toc}{\protect\thispagestyle{empty}}
%\pagestyle{empty}
%\cleardoublepage
\newpage
%\pagenumberingstyle{plain}
\setcounter{page}{1} %reset the page counter
\chapter{Introduction}
\thispagestyle{plain}
% more intro to gf. dependently typed...high-level formalism
Grammatical Framework \cite{gfbok} is a dependently typed grammar formalism.
%with facilities for multilinguality. 
It is based on Martin-Löf type theory which allows
reasoning within the programming language.
GF has strong support for multilinguality and has so far been used successfully for 
controlled languages \cite{cnl}, while recent experiments have showed
that it is also possible to use the framework for parsing free language \cite{patent}.
% >> although the success rate is going steeply down when using uncontrolled nl. Better with annotated.
% --> sentence about connection GF - grammar based parser.
%This project focuses on implementing a wide-coverage Swedish grammar, which
%describes a parser accepting all sentences that can be expressed by the grammar.
Parsing, that is the task of automatically identifying morphological and
syntactical structures, is receiving increasing interest, especially
considering the steadily growing amount of data available online. 
So far there is no freely available grammar-driven parser that gives a deep
analysis for Swedish. The fact that %the existing rule-based Swedish
a parser or grammar is not freely available does not only restrict its
usage but also its possibilities of being further developed. Our
goal is
to create an open-source Swedish grammar from which we derive a parser
accepting all sentences described by the given rules.
As a freely available resource, it can be continuously enhanced and
made use of in other projects.\\

@@

Apart from these restrictions, noun phrases containing reflexive pronouns
may be used as any other \verb-NP-. They may be conjoined 
(\ref{ex:ReflConj} a,b) and used with other determiners (\ref{ex:ReflPredet}).
In the standard GF analysis, which is preformed bottom-up starting from the
%TODO verify
POS-tags, information about semantic roles are given by the functions, not by the
categories. That is,
we know that the first argument of the function \verb-PredVP- acts as the
subject, but
the noun phrase itself does not carry information about its semantic role.
Until it is given as an argument to a clause level function, no difference
is made between subject and object noun phrases.
For this reason,
the formalization of reflexive pronouns required the use of a different analysis.
%and used as genitives (sentence \ref{ex:ReflGen}) as such.
%Could be combined with 'sig', a ReflNP without another noun.
In short, what is wanted can be summarized as follows:\\
%\begin{figure}[h]
\begin{enumerate}
\item{
Construct a noun phrase from a common noun phrase \\
katt $\rightarrow$ sin katt} % \\
\item{Modify it like a \verb|NP|\\
sina katter $\rightarrow$ alla sina katter}
\item{Use it only as object\\
sin katt $\rightarrow$ han såg sin katt }
\end{enumerate}
%\caption{Rules for noun phrases with reflexive possessive pronouns}
%\label{fig:reflrestrict}
%\end{figure}
